---
Title: "predictimg velocity of prediction"
Author: "Hari Chandana"
Date: "July 16, 2018"
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->
#description
This package predicts the velocity of progression given training data

## 1. Dataset
KDD dataset is used 
https://pslcdatashop.web.cmu.edu/KDDCup/rules.jsp

## 2.Model built: Decision tree:
Decision trees implicitly perform feature selection
To overcome scale differences between parameters
Decision trees are also not sensitive to outliers since the splitting happens based on proportion of samples within the split ranges and not on absolute values. 

## 3. Functions in package
The kdd2 package consists of three functions
###3a ppify.data
This function preprocesses the kdd data
```{r}

ppify.data = function(raw.data = "/usr/local/src/app/modified_kddv2.csv",
                      preprocessed.file = "/usr/local/src/app/kddv2.csv"){

  #raw.data = read_csv("E:/bitwise/modified_kddv2.csv")
 
  #loading required libraries
  require(pracma)
  require(dplyr)
  require(jsonlite)
  
  # raw.data can either be csv file or json input data.
  newdata = if(is.character(raw.data) && file.exists(raw.data)){
    read.csv(raw.data, sep = ',', header = T)
  }	else{
    as.data.frame(raw.data)
  }
  
  #assigning newdata to kdd
  kdd<-newdata

  #converting follwing variables to factor variables
  kdd$Problem.Name=as.factor(kdd$Problem.Name)
  kdd$Anon.Student.Id=as.factor(kdd$Anon.Student.Id)
  
  #creating new attributes
  oppskills=kdd$Opportunity.SubSkills.+kdd$Opportunity.KTracedSkills.
  countskills=kdd$Count.KTracedSkills.+kdd$Count.SubSkills.
  
  #binding them to the dataframe
  kdd<-cbind(kdd,oppskills)
  kdd<-cbind(kdd,countskills)
  
  #selecting required attributes
  kdd<- kdd[c("Anon.Student.Id","Subject", "Step.Duration..sec.","Correct.First.Attempt","oppskills","countskills","Hints","Problem.View")]

  #summary(newdata$oppskills)

  #creating new attribute diff by dividing it into intervals based on quartile
  diff <- cut(kdd$oppskills,c(0,20,92,179.6,250),include.lowest = TRUE,labels=c("1","2","3","4"))
  diff<-as.numeric(diff)
  kdd<-cbind(kdd,diff)
 
  #Percentage of Correct first attemps upto now (excluding the current row)
  kdd$EWMA_assess = 0 
  kdd$MA_session = 0
  # Velocity of Progression of a student
  kdd$vop = 0
  genpop_mean<-mean(kdd$Step.Duration..sec.)
  

  studentid <- array(unique(kdd$Anon.Student.Id))
  
  # Finding the exponentially weighted moving average of the assessment score
  for (num_i in 1:length(studentid)){
      # Assigning the value of window based on condition
      window = 3
      min_window=2
      if(nrow(kdd[kdd$Anon.Student.Id == studentid[num_i], ]) > window &
         nrow(kdd[kdd$Anon.Student.Id == studentid[num_i], ]) < min_window){
        window = nrow(kdd[kdd$Anon.Student.Id == studentid[num_i] , ])
      } else if(nrow(kdd[kdd$Anon.Student.Id == studentid[num_i] , ]) <= min_window){
        window = 1 # Single Entry or window size data length Flag
      }
      if (window > 1) {
        kdd[kdd$Anon.Student.Id == studentid[num_i] ,
                c("Step.Duration..sec.", "Correct.First.Attempt", "MA_session", "EWMA_assess")] <-
          within(kdd[kdd$Anon.Student.Id == studentid[num_i] ,
                         c("Step.Duration..sec.", "Correct.First.Attempt", "MA_session", "EWMA_assess")],
                 { EWMA_assess <- pracma::movavg(Correct.First.Attempt, window, type='e'); # calculate  EWMA for each row from the past 3 rows
               MA_session <- dplyr::cummean(Step.Duration..sec.)  }    )
      } else {
        kdd[kdd$Anon.Student.Id == studentid[num_i] ,
                c("Step.Duration..sec.", "Correct.First.Attempt", "MA_session", "EWMA_assess")] <-
          within(kdd[kdd$Anon.Student.Id == studentid[num_i],
                         c("Step.Duration..sec.", "Correct.First.Attempt", "MA_session", "EWMA_assess")],
                 { EWMA_assess <- mean(Correct.First.Attempt); # For single entry, we pass them as it is   same as above
                MA_session <- mean(Step.Duration..sec.) })
     }

}


```
This function is used to preprocess the kdd dataset.
We assigned value to window based on condition and then defined new attributes EWMA_assess(Percentage of correct first attemps till now,wxcluding the correct row),MA_session.
We defined the value of Vop(velocity of progression)
This function saves the preprocessed file

###3b buildmodel
```{r}
buildmodel=function(save.model="/usr/local/src/app/kddmodel.rds"){
  #loading required libraries
  #library(reshape2)
  library(jsonlite)
  library(mongolite)
  library(data.table)
  library(tidyr)
  #creation of mongo connection object that connects to collection kdd2 in database chandana_test
  conn <- mongo(collection = "kdd2", db = "chandana_test", url = "mongodb://admin:bitwise2017@13.78.179.80:27017/admin",verbose = TRUE)
  alldata<-conn$find('{}')#empty querying to import the data 
  dt<-data.table(alldata)
  #creating a vector with attribute names
  cols <- c("Hints","countskills","diff","MA_session","EWMA_assess","vop")
  #using separate_rows to split the data
  dt<-tidyr::separate_rows(dt,cols[1:6])
  dt1<-dt[!dt$Hints == "", ]#cleaning the data
  dt2<-dt1[!dt1$Hints == "c", ]
  library(rpart)
  
  #building the decision regression tree
  model2<- rpart(vop~ Hints+countskills+diff+MA_session+EWMA_assess, data = dt2,method = "anova")
  saveRDS(model2,save.model )#save the model
}

```
This function imports the data from mongodb using mongo connection object conn and then builds the decision regression tree
Then, it saves the model

###3c predict.data
```{r}

predict.data = function(input,save.model="/usr/local/src/app/kddmodel.rds"){

  # raw.data can either be csv file or json input data.
  newdata = if(is.character(input) && file.exists(input)){
    read.csv(input, sep = ',', header = T)
  }       else{
    as.data.frame(input)
  }
  model2=readRDS(file=save.model)#load the model
  predicttrain2=predict(model2,newdata)
  return(predicttrain2)
}

```
This function loads the saved model and predicts the outputs
