---
Title: "Student Performance prediction"
Author: "Ashwin V "
Date: "July 16, 2018"
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->
#description
This package predicts the performance and complexity of a problem given input data

## 1. Dataset
The dataset used is from the KDD Cup 2010 Education Dataset https://pslcdatashop.web.cmu.edu/KDDCup/rules.jsp

## 2.Models
This packages is built on 2 models
###2a: Decision tree:
Decision trees implicitly perform feature selection to overcome scale differences between parameters Decision trees are also not sensitive to outliers since the splitting happens based on proportion of samples within the split ranges and not on absolute values. 
###2b: K-Means Cluster:

## 3. Functions in package
The kdd package consists of three functions
###3a build_tree
This function builds the regression tree model and saves the model in the working directory
```{r}

build_tree = function(){
  df = mongo$find('{}')
  l1 <- sapply(df$Hints, length)
  xx<- unlist(rep(df$Anon_Student_Id, l1))
  sample <- df[,-1] 
  df.new <- data.frame(Anon_Student_Id = xx)
  for(j in 1:ncol(sample)){
    x <- data.frame(unlist(sample[,j]))
    df.new <- cbind(df.new,x)
  }
  colnames(df.new) <- names(df)
  df.new <- na.omit(df.new)
  
  decisiontree <- rpart(Percent_First_Attempts ~ No_Step + Step_Duration__sec_ +Hints, data=df.new, method="anova") 
  saveRDS(decisiontree,file ="/datadrive/TEMP/ashwin/performance_modelinfo.RData" )
}


```
This function is used to preprocess the dataset retrieved form mongodb. A regression tree model is built based on the number of hints, total duration, number of steps and is used to predict the percentage of correct first attempts. This function saves the model in working directory.

###3b buildmodel
```{r}
predict_performance = function(input){
  df = mongo$find('{}')
  l1 <- sapply(df$Hints, length)
  xx<- unlist(rep(df$Anon_Student_Id, l1))
  sample <- df[,-1] 
  df.new <- data.frame(Anon_Student_Id = xx)
  for(j in 1:ncol(sample)){
    x <- data.frame(unlist(sample[,j]))
    df.new <- cbind(df.new,x)
  }
  colnames(df.new) <- names(df)
  df.new <- na.omit(df.new)
  df.new <- rbind.fill(df.new,input)
  
  fit <- kmeans(df.new[,c(4,6,8,11,12)], 3) 
  Fit <-data.frame(unlist(fit$cluster))
  colnames(Fit) <- c("category")
  course <- cbind(df.new,Fit) 
  course$Diff_level[course$category==1]="Easy"
  course$Diff_level[course$category==2]="Hard"
  course$Diff_level[course$category==3]="Medium"
  print("Difficulty of the problem = ")
  print(tail(course$Diff_level,1))
  
  performance=readRDS("/datadrive/TEMP/ashwin/performance_modelinfo.RData")
  print("Performance = ")
  predict(performance,input, method = "anova")
}

```
This function imports the data from mongodb using mongo connection object conn and then builds the k-means cluster to predict the difficulty of a problem. Siilarly when an input is given, the student performance and difficulty of the problem is displayed.


